---
title: "Application of the Central Limit Theorem to the Exponential Distribution"
author: "Nathan Buesgens"
output: html_document
---

This report uses simulated samples from an exponential distribution to demonstrate the
principles of the Central Limit Theorem. Sample mean, variance, and distribution will approximate
population mean, variance and distribution as sample size grows large. Regardless of the population distribution, the Central
Limit Theorem states that sample means will be normally distributed about the population mean with variance inversely proportional to the sample size.

```{r, echo=FALSE}
lambda = 0.2
population_mean = 1/lambda
population_sd = 1/lambda
```

## Simulations
The following data is a simulation of a sample of size n = 1000 from a population that has an exponential distribution:
```{r}
n = 1000
single_sample = rexp(n, lambda)
```
This is to be compared to data that is a simulation of m = 1000 samples of size n = 40 from a population that has an exponential distribution:
```{r}
m = 1000
n = 40
fourty_sample_means = NULL
for (i in 1 : m) fourty_sample_means = c(fourty_sample_means, mean(rexp(n, lambda)))
```

## Sample Mean vs Theoretical Mean
The sample mean of a single sample approximates the population mean as n grows large by
the Law of Large Numbers. Here the sample mean is blue and the population mean in red:

```{r, fig.width=5, fig.height=2.5, echo=FALSE}
single_sample_mean = mean(single_sample)
hist(single_sample, main="Mean of a Single Sample", xlab="index")
abline(v = single_sample_mean, col = "blue", lwd = 2)
abline(v = population_mean, col = "red", lwd = 2)
```

The mean of sample means are normally distributed by the Central Limit Theorem, centered
around the population mean:

```{r, fig.width=5, fig.height=2.5, echo=FALSE}
mean_of_means = mean(fourty_sample_means)
hist(fourty_sample_means, main="Mean of Sample Means", xlab="mean")
abline(v = mean_of_means, col = "blue", lwd = 2)
abline(v = population_mean, col = "red", lwd = 2)
```

## Sample Variance vs Theoretical Variance

The population variance can be approximated by the sample variance:

```{r, fig.width=5, fig.height=2.5, echo=FALSE}
hist(single_sample, main="Sample Variance", xlab="index")
abline(v=single_sample_mean, lwd=2)
abline(v=single_sample_mean + sd(single_sample), col="blue")
abline(v=single_sample_mean + population_sd, col="red")
```

The variance of sample means is $\sigma^2/n$ by the Central Limit Theorem. The population variance
can be approximated by the sample variance. The blue line is one standard deviation from the mean approximated by a single sample variance. The red line is one standard deviation calculated using the population variance:

```{r, fig.width=5, fig.height=2.5, echo=FALSE}
mean_sd = sd(rexp(n, lambda))/sqrt(n)
hist(fourty_sample_means, main="Variance of Sample Means", xlab="mean")
abline(v = mean_of_means, lwd = 2)
abline(v = mean_of_means + mean_sd, col = "blue", lwd = 2)
abline(v = population_mean + population_sd/sqrt(n), col = "red", lwd = 2)
```

## Distribution

Random samples from a population will have a distribution similar to the population while
by the Central Limit Theorem means of samples will be distributed normally regardless of the population distribution:

```{r, fig.width=5, fig.height=2.5, echo=FALSE}
hist(fourty_sample_means, main="Distribution of Sample Means", xlab="mean", prob=TRUE)
lines(seq(0, 8, length = 100), dnorm(seq(0, 8, length = 100), population_mean, population_sd/sqrt(n)), col = grey(.8), lwd = 3)
```


